---
title: "IV3 Local Average Treatment Effects"
subtitle: "ECON 474 Econometrics of Policy Evaluation"
output: 
  html_document:
    toc: yes
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# [Lab] Card (1993)
********************************************************************************

Today we have a short lab to practice using the `feols()` function and to think about the local part of local average treatment effects.
We will be using data from Card (1993) "Using Geographic Variation in College Proximity to Estimate the Return to Schooling" and doing what the title implies.
Specifically, we will be instrumenting the endogenous "years of schooling" variable with an indicator variable for an accredited four-year university.


## Setup
**********************************************************************

1. Load in the packages: `tidyverse` and `fixest`
2. Load in the data `card.csv`
3. Print the head of the data


```{r loading}

```

## Replication 
**********************************************************************

Produce an `etable()` of the biased OLS regression, the reduced form, the first stage, and the structural model using heteroskedastic robust standard errors and the controls listed below.

Variables:

- y: log wage (`lwage`)
- d: years of schooling (`educ`)
- z: college-in-the-county (`nearc`)
- x: a quadratic in years of experience (`exper`), an indicator for being black (`black`), an indicator for living in the south (`south`), and an indicator for living in an MSA (`smsa`)

Within your `etable()`

- display the number of observations, the $R^2$ and the F-statistic
- label your regressions appropriately using `header`
- `order` the variables to have the `Intercept`, years of schooling, and college-in-the-county come first in that order.

```{r etable}

```


Education is endogenous because of omitted variable bias (i.e. ability). 
We would expect the OLS estimate to be biased larger than the "true" estimate.
Does the 2SLS least squares return this result (*hint: it doesn't*)?
If not, what could be an economic (not statistical) explanation to the results?

--your answer here--