---
title: "DiD1 2x2 Difference-in-Differences"
subtitle: "ECON 474 Econometrics of Policy Evaluation"
output: 
  html_document:
    toc: yes
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Here we are demonstrating the $2 \times 2$ difference-in-differences.
What makes it $2 \times 2$ is that we will be running a regression using a variable for the two groups (treatment and control) and a variable for the two time periods (pre-treatment and post-treatment).

To begin, load in our two favorite packages `tidyverse` and `fixest` along with the data.
The data are tiny, so we can also print it.


```{r loading}
library(tidyverse)
library(fixest)

banks = read_csv("C:/Users/johnj/Dropbox/Julian/Academia/UIUC/Semester 12 Spring 2023/ECON474sp23/Data/DiD1 Mississippi experiment.csv")

banks
```

Here we can see that we have data over 6 years.
The variables `6th district` and `8th district` are the number of banks operating in each Federal Reserve district.
Recall that district 6 is the Atlanta district (treated) and that distrcit 8 is the St. Louis district (control).

# Parallel Trends
********************************************************************************

Parallel trends means that whatever happens to the control group after treatment is exactly what would have happened to the treatment group if they were not treated.
We can hard code this into the data for demonstration purposes of what we require.

To be clear, this **does not provide evidence** that the parallel trends assumption is satisfied.
It is only showing what is required for the assumption to be satisfied.


```{r parllel trends df}
df_plot = banks %>%
  mutate(diff_d8_d6 = `8th district`[year == 1930] - `6th district`[year == 1930],
         counterfactual = if_else(year >= 1930,
                                  `8th district` - diff_d8_d6,
                                  `6th district`)) %>%
  select(-diff_d8_d6) %>%
  pivot_longer(cols = c(`6th district`, `8th district`, counterfactual),
               names_to = 'district',
               values_to = "n_banks")

```

Now, let's plot this to see what we the unobservable counterfactual should look like.

```{r parllel trends plot}
ggplot(df_plot, aes(x = year, y = n_banks, color = district, linetype = district)) +
  geom_line() +
  scale_linetype_manual(values = c(rep("solid", 2), "dashed")) +
  labs(x = "Year", y = "Number of Banks in Business") +
  geom_vline(xintercept = 1930)
```


# 2x2 Difference-in-Differences
********************************************************************************

To run our first diff-in-diff, we need to take the `df_plot` and remove the made-up counterfactual outcome.
When creating a new object, we will also create the two variables that we need to run the regression: `treat` and `post`, which are self-explanatory.


```{r df reg}
df_reg = df_plot %>%
  filter(district != "counterfactual") %>%
  mutate(treat = (district == "6th district")*1,
         post = (year > 1930)*1)
```

Time to run the 2x2 regression!
But first, since we have multiple pre-treatment observations and multiple post-treatment observations for each unit, we need to include fixed effects.
However, I want to show you how to run a standard diff-in-diff, so we will estimate one regression without fixed effects and another with.
Take a look at which coefficients are reported for the second regression.
Is this a problem?


```{r 2x2}
fit1 = feols(n_banks ~ treat*post, data = df_reg)
fit2 = feols(n_banks ~ treat*post | district, data = df_reg)

etable(fit1, fit2,
       se.below = TRUE)
```


To conclude, I want to show you visually what these regressions are actually estimating.
The short answer is averages.
Try to think about how the plotted averages map into the potential outcomes framework.

```{r 2x2 plot}
ggplot(df_reg, aes(x = year, y = n_banks, color = district)) +
  geom_line() +
  labs(x = "Year", y = "Number of Banks in Business", color = "") +
  geom_smooth(method = "lm",
              formula = y ~ I(x > 1930),
              se = FALSE,
              linetype = "dashed")

```



