---
title: "IV4 Treatment on the Treated"
subtitle: "ECON 474 Econometrics of Policy Evaluation"
output: 
  html_document:
    toc: yes
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# The Oregon Heatlh Insurance Experiment
********************************************************************************

The Oregon health insurance experiment (HIE) provides an excellent demonstration of using instrumental variables in a randomized experiment when there is one-sided non-compliance. 
We are going to replicate part of [Finkelstein et a. 2012](https://academic.oup.com/qje/article/127/3/1057/1923446) to demonstrate the relationship between intent-to-treat (ITT) and treatment-on-the-treated (TOT) with the three causal equations of instrumental variables.




## Preliminaries
**********************************************************************

As always, we are going to begin with the standard loading in of a few packages and the data.
Don't print the `head()` here because this data has a plethora (over 500) of variables.

```{r loading}
library(haven)
library(tidyverse)
library(fixest)

hie = read_dta('C:/Users/JJPWade/Dropbox/Julian/Academia/UIUC/Semester 12 Spring 2023/ECON474sp23/Data/IV4 oregonhie_data_for_analysis.dta')

```

Finkelstein et al. 2012 use two sources of data: one that is from administrative government sources and another that is a survey that some of the individuals in the administrative data responded to.
The authors have helped us identify which variables should be used as controls for data source.
The administrative data controls begin with either a `lll` or an `nnn`.


```{r admin controls}
vrbls = names(hie)
# vrbls[grepl('lll', vrbls)]
# vrbls[grepl('nnn', vrbls)]
vrbls[grepl('lll|nnn', vrbls)]
```

The `lll` prefixes indicate which lottery round are individuals obtained access to the Oregon state Medicaid.
The `nnn` prefixes are indicators for number of people in household.

The survey data controls begin with a `ddd`.

```{r survey controls}
vrbls[grepl('ddd', vrbls)]
```

There are three variables that have the `ddd` prefix:

1. `draw_lot` - survey wave
2. `num_hh` - number of people in household
3. `raXnum` the interaction of `draw_lot` and `num_hh`


### Levelling up our regression coding
************************************************************

Throughout this replication, we are going to be estimating many regressions that are identical but with *multiple outcomes*.
Fortunately, `feols()` handles "multiple outcome" regression quite well by inputting the outcome variables without quotes into `c()`.


## Table 3 - First Stage Estimates 
**********************************************************************



We are going to begin with the administrative data first.
Recall that we are going to put all of the outcome variables into `c()` within the `feols()` `formula`.
We could type out all of the control variables, or we could instead use our variable selection using `grepl()` from above inside of `paste()` to speed up by copying and pasting into the `formula`.

Something that you should include in all of your regression tables is the average outcome value for the control group.
The reason why this is important is so we can compare it with the regression coefficients to determine the effect size and economic significance.
For example, if cancer rates go up by 100%, but only 1 out of 1,000,000,000 got cancer in the control group, then we know the effect size does not have economic significance.

As we are going to run our first regression, something important to note that Medicaid is applied *not to individuals* but to *all individuals within a household*. 
Because our data are on individuals, we need to cluster the data to the treatment level: households.


```{r first stage admin}
paste(vrbls[grepl('nnn|lll', vrbls)], collapse = ' + ')

fs_fits_hdd = 
  feols(c(ohp_all_ever_admin, ohp_std_ever_admin, ohp_all_mo_admin, ohp_all_end_admin) ~ treatment + llldraw_lot_2 + llldraw_lot_3 + llldraw_lot_4 + llldraw_lot_5 + llldraw_lot_6 + llldraw_lot_7 + llldraw_lot_8 + nnnnumhh_li_2 + nnnnumhh_li_3,
      data = hie,
        cluster = ~ household_id)

fs_control_means_hdd = hie %>%
  filter(treatment == 0) %>%
  summarise_at(vars(ohp_all_ever_admin, ohp_std_ever_admin, ohp_all_mo_admin, ohp_all_end_admin), mean) %>%
  unlist()
```

Now we are going to do the same thing, but for the survey data.
Because not everyone in the experiment responded to the survey, the authors created survey weights to make those that did respond represent the experiment population.
We need to adjust our `feols()` and control group mean calculations accordingly.

```{r first stage survey}
paste(vrbls[grepl('ddd', vrbls)], collapse = ' + ')

fs_fits_12m =
  feols(c(ohp_all_ever_survey, ohp_std_ever_survey, ohp_all_mo_survey, ohp_all_end_survey) ~ treatment + ddddraw_sur_2 + ddddraw_sur_3 + ddddraw_sur_4 + ddddraw_sur_5 + ddddraw_sur_6 + ddddraw_sur_7 + dddnumhh_li_2 + dddnumhh_li_3 + ddddraXnum_2_2 + ddddraXnum_2_3 + ddddraXnum_3_2 + ddddraXnum_3_3 + ddddraXnum_4_2 + ddddraXnum_5_2 + ddddraXnum_6_2 + ddddraXnum_7_2,
        data = hie,
        cluster = ~ household_id,
        subset = ~ sample_12m_resp == 1,
        weights = ~ weight_12m)
  
fs_control_means_12m =
  hie %>%
  filter(treatment == 0 & 
           sample_12m_resp == 1) %>%
  summarise_at(vars(ohp_all_ever_survey, ohp_std_ever_survey, ohp_all_mo_survey, ohp_all_end_survey),
               weighted.mean,
               w = hie$weight_12m[hie$treatment == 0 & 
           hie$sample_12m_resp == 1]) %>%
  unlist()
  

```

Great! Now we can put our results into an `etable()` to present our results.
We are going to omit the coefficients for the controls to be parsimonious.
Remember to always describe which variables are in your regression whenever presenting your results, especially if you are choosing to not display them.

```{r first stage 1st 4 rows}
etable(fs_fits_hdd, fs_fits_12m,
       se.below = TRUE,
       keep = 'treatment',
       fitstat = ~ n + r2 + f,
       extralines = list('_^Control Mean Outcome' = c(fs_control_means_hdd, fs_control_means_12m)))

```



## Table 5 - Healthcare Utilization
**********************************************************************

Unfortunately for us (but fortunately for the respondents) the administrative is not publicly disclosed to maintain confidentiality.
That means we can only analyze individuals who responded to the survey, but this is sufficient to demonstrate ITT and TOT.

Table 5 shows both the ITT and LATE (TOT) for several healthcare utilization outcomes.
We are only going to show a few of the outcomes reported to demonstrate the process.
Nonetheless, because ITT is estimated from the reduced form and TOT is estimated from the structural equation, we need to run both of these regressions inside of our for loop.


```{r tab5}
# ITT - Reduced form
fit5_itt = feols(c(rx_any_12m, rx_num_mod_12m, doc_any_12m, doc_num_mod_12m) ~ treatment + ddddraw_sur_2 + ddddraw_sur_3 + ddddraw_sur_4 + ddddraw_sur_5 + ddddraw_sur_6 + ddddraw_sur_7 + dddnumhh_li_2 + dddnumhh_li_3 + ddddraXnum_2_2 + ddddraXnum_2_3 + ddddraXnum_3_2 + ddddraXnum_3_3 + ddddraXnum_4_2 + ddddraXnum_5_2 + ddddraXnum_6_2 + ddddraXnum_7_2,
        data = hie,
        cluster = ~ household_id,
        subset = ~ sample_12m_resp == 1,
        weights = ~ weight_12m)

# TOT - Structural equation
fit5_tot = feols(c(rx_any_12m, rx_num_mod_12m, doc_any_12m, doc_num_mod_12m) ~  ddddraw_sur_2 + ddddraw_sur_3 + ddddraw_sur_4 + ddddraw_sur_5 + ddddraw_sur_6 + ddddraw_sur_7 + dddnumhh_li_2 + dddnumhh_li_3 + ddddraXnum_2_2 + ddddraXnum_2_3 + ddddraXnum_3_2 + ddddraXnum_3_3 + ddddraXnum_4_2 + ddddraXnum_5_2 + ddddraXnum_6_2 + ddddraXnum_7_2 | ohp_all_ever_survey ~ treatment,
        data = hie,
        cluster = ~ household_id,
        subset = ~ sample_12m_resp == 1,
        weights = ~ weight_12m)

# control group average outcome
fit5_control_means = 
  hie %>%
  filter(treatment == 0 & 
           sample_12m_resp == 1) %>%
  summarise_at(vars(rx_any_12m, rx_num_mod_12m, doc_any_12m, doc_num_mod_12m),
               weighted.mean,
               w = hie$weight_12m[hie$treatment == 0 & 
           hie$sample_12m_resp == 1],
           na.rm = TRUE) %>%
  unlist()

```

We have ran two types of regressions, which means we need to report two variables: the instrument `treatment` and the endogenous variable `ohp_all_ever_survey`.


```{r tab5 rep}
etable(fit5_itt, fit5_tot,
       se.below = TRUE,
       keep = c('treatment', 'ohp_all_ever_survey'),
       fitstat = ~ n + r2 + f,
       extralines = list('_^Control Mean Outcome' = fit5_control_means))
```



## Table 9 - Health
**********************************************************************

There are three classes of outcomes Finkelstein et al. look at:

1. utilization
2. finances
3. health

My prior is that if people have access to health insurance then they will utilize healthcare more.
Their results on utilization don't surprise me.
While we are not replicate the finance results, they find that people with health insurance experience less medical related financial burdens. 
Again, that is what I would expect.

The interesting part of provided access to health insurance is does it improve health.
I expect that going to the doctor when sick/injured would be helpful, but I also expect that simply having health insurance does not change our habits (eating/exercise/stress-management) that contribute to health. 
Let's take a look at a few of the outcomes in table 9 to see if people truly did become healthier according to their self-reports.


```{r tab9}
fit9_itt = feols(c(health_genflip_bin_12m, notbaddays_phys_12m, notbaddays_ment_12m) ~ treatment + ddddraw_sur_2 + ddddraw_sur_3 + ddddraw_sur_4 + ddddraw_sur_5 + ddddraw_sur_6 + ddddraw_sur_7 + dddnumhh_li_2 + dddnumhh_li_3 + ddddraXnum_2_2 + ddddraXnum_2_3 + ddddraXnum_3_2 + ddddraXnum_3_3 + ddddraXnum_4_2 + ddddraXnum_5_2 + ddddraXnum_6_2 + ddddraXnum_7_2,
                 data = hie,
                 cluster = ~ household_id,
                 subset = ~ sample_12m_resp == 1,
                 weights = ~ weight_12m)

# TOT - Structural equation
fit9_tot = feols(c(health_genflip_bin_12m, notbaddays_phys_12m, notbaddays_ment_12m) ~  ddddraw_sur_2 + ddddraw_sur_3 + ddddraw_sur_4 + ddddraw_sur_5 + ddddraw_sur_6 + ddddraw_sur_7 + dddnumhh_li_2 + dddnumhh_li_3 + ddddraXnum_2_2 + ddddraXnum_2_3 + ddddraXnum_3_2 + ddddraXnum_3_3 + ddddraXnum_4_2 + ddddraXnum_5_2 + ddddraXnum_6_2 + ddddraXnum_7_2 | ohp_all_ever_survey ~ treatment,
                 data = hie,
                 cluster = ~ household_id,
                 subset = ~ sample_12m_resp == 1,
                 weights = ~ weight_12m)

# control group average outcome
fit9_control_means = 
  hie %>%
  filter(treatment == 0 & 
           sample_12m_resp == 1) %>%
  summarise_at(vars(health_genflip_bin_12m, notbaddays_phys_12m, notbaddays_ment_12m),
               weighted.mean,
               w = hie$weight_12m[hie$treatment == 0 & 
                                    hie$sample_12m_resp == 1],
               na.rm = TRUE) %>%
  unlist()



etable(fit9_itt, fit9_tot,
       se.below = TRUE,
       keep = c('treatment', 'ohp_all_ever_survey'),
       fitstat = ~ n + r2 + f,
       extralines = list('_^Control Mean Outcome' = fit9_control_means))

```


# Conclusion
********************************************************************************

We wrap up our conversation on instrumental variables tying back into the "gold standard" of randomized experiments. 
Even though it may be the conceptual ideal, there are still that can occur that can bias average treatment effect estimates.
For example, non-compliance changes $\hat{\delta}$ from ATE to ITT. 
The closest we can get to ATE is the TOT (also called ATT), but to do so we need to use instrumental variables.

Instrumental variables can be hard to identify because there are many (5) assumptions that need to be satisfied, where only the inclusion restriction is testable. 
However, it is an incredibly powerful method because of how it can still estimate a LATE in the presence of bias from endogeniety sources like:

- omitted variable bias
- choice variables
- multiple equations (supply and demand)
- non-compliance